The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-03 analysis/paper/paper.qmd
Add script to compute word embeddings

The `tf-idf` statistic shows high values for the names of thesis locations that tend to be high-frequency thesis-specific words. As such, it shows words that are the most important to one document in a collection of documents [@silge2017].

>>>>

```{r word-embeddings}

tidy_emu_theses <- emu_theses |> 
  select(title, text_clean) |> 
  unnest_tokens(word, text_clean) |> 
  add_count(word) |> 
  filter(n >= 50) |> 
  select(-n)

nested_words <- tidy_emu_theses |> 
  nest(words = c(word))

nested_words

slide_windows <- function(tbl, window_size) {
  skipgrams <- slider::slide(
    tbl,
    ~.x,
    .after = window_size - 1,
    .step = 1,
    .complete = TRUE
  )

  safe_mutate <- safely(mutate)

  out <- map2(skipgrams,
              1:length(skipgrams),
              ~ safe_mutate(.x, window_id = .y))

  out |> 
    transpose() |> 
    pluck("result") |> 
    compact() |> 
    bind_rows()
}

library(widyr)
library(furrr)

plan(multisession)

tidy_pmi <- nested_words |> 
  mutate(words = future_map(words, slide_windows, 10L)) |> 
  unnest(words) |> 
  unite(window_id, title, window_id) |> 
  pairwise_pmi(word, window_id)

tidy_pmi

tidy_word_vectors <- tidy_pmi |> 
  widely_svd(
    item1, item2, pmi,
    nv = 100, maxit = 1000
  )

tidy_word_vectors

nearest_neighbors <- function(df, token) {
  df |> 
    widely(
      ~ {
        y <- .[rep(token, nrow(.)), ]
        res <- rowSums(. * y) /
          (sqrt(rowSums(. ^ 2)) * sqrt(sum(.[token, ] ^ 2)))

        matrix(res, ncol = 1, dimnames = list(x = names(res)))
      },
      sort = TRUE
    )(item1, dimension, value) |> 
    select(-item2)
}

tidy_word_vectors |> 
  nearest_neighbors("resilience")

tidy_word_vectors |> 
  filter(dimension <= 24) |> 
  group_by(dimension) |> 
  top_n(12, abs(value)) |> 
  ungroup() |> 
  mutate(item1 = reorder_within(item1, value, dimension)) |> 
  ggplot(aes(item1, value, fill = dimension)) +
  geom_col(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~dimension, scales = "free_y", ncol = 4) +
  scale_x_reordered() +
  coord_flip() +
  labs(
    x = NULL,
    y = "Value",
    title = "First 24 principal components for text of theses",
    subtitle = "Top words contributing to the components that explain the most variation"
  )

word_matrix <- tidy_emu_theses |> 
  count(title, word) |> 
  cast_sparse(title, word, n)

embedding_matrix <- tidy_word_vectors |> 
  cast_sparse(item1, dimension, value)

doc_matrix <- word_matrix %*% embedding_matrix

dim(doc_matrix)

doc_matrix
```
